Graph Topological Aspects of Granger Causal Network Learning

### 2019-06-30 paper/paper.tex
Some descriptions for figures and minor updates.

\begin{theorem}[Pairwise Recovery]
  \label{thm:tree_recovery}
  If the Granger-causality graph $\gcg$ for persistent process $x(t)$
  is a strongly causal DAG then $\gcg$ can be inferred from pairwise
  causality tests.  The procedure can be carried out, assuming
  we have an oracle for pairwise causality, via Algorithm
  (\ref{alg:pwgr}).

@@

% \subsection{Pairwise Minimum Error Spanning Trees}
% Inspired by theorem \ref{thm:tree_recovery}, we propose the following
% structure learning heuristic.

@@

where we are choosing $\lambda$, the regularization parameter, via the BIC.

@@

\begin{figure}
  \centering
  \caption{Simulation Results}
  \label{fig:simulation_results}

  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
    estimation.  The MCC is measured on the left axis in blue with
    circle markers, and the relative log mean-squared prediction error
    $\frac{\ln\ \tr \widehat{\Sigma}_v}{\ln\ \tr \Sigma_v}$ on the
    right axis in red with triangle markers.  We use $q$ to denote the
    edge probability in a random Erdos-Reny graph (modified to form a
    DAG).  Solid lines indicate ad-hoc regression estimates for
    visualization purposes only.

    Matthew's Correlation Coefficient (MCC)
    \cite{matthews1975comparison} synthesizes the confusion matrix
    into a single score appropriate for unbalanced labels and is
    calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
    performance, $0$ being the performance of random guessing, and
    $-1$ being perfectly opposed.
  }

@@

  \begin{subfigure}[b]{0.45\textwidth}
    \caption{Fixed $T$, increasing $n$ (SCG)}
    \label{fig:simulation_results_scaling}
    \includegraphics[width=\linewidth]{increasing_n_simulation.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
    \label{fig:simulation_results_dense}
    \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}
  \end{subfigure}

  \begin{subfigure}[b]{0.45\textwidth}
    \caption{MCC Comparison for $T \le 100$}
    \label{fig:mcc_comparison}
    \includegraphics[width=\linewidth]{new_mcc_comparison000.pdf}
  \end{subfigure}
\end{figure}

>>>>

\begin{theorem}[Pairwise Recovery]
  \label{thm:scg_recovery}
  If the Granger-causality graph $\gcg$ for persistent process $x(t)$
  is a strongly causal DAG then $\gcg$ can be inferred from pairwise
  causality tests.  The procedure can be carried out, assuming
  we have an oracle for pairwise causality, via Algorithm
  (\ref{alg:pwgr}).

@@

% \subsection{Pairwise Minimum Error Spanning Trees}
% Inspired by theorem \ref{scg_recovery}, we propose the following
% structure learning heuristic.

@@

where we are choosing $\lambda$, the regularization parameters, via the BIC.

@@

\begin{remark}[Graph Topologies]
  We depict in Figure \ref{fig:random_graph_topologies} the topologies
  of random graphs used in our empirical evaluation.  For values of
  $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
  a topology which is, at least qualitatively, close to the SCG.  As
  the value of $q$ increases, the random graphs deviate farther from
  the SCG topology, and we therefore expect the LASSO to outperform
  PWGC for larger values of $q$.  This can be observed in Figure
  \ref{fig:simulation_results_dense}, at least where performance is
  measured by the support recovery (i.e. via the MCC).
\end{remark}

\begin{remark}[MCC as a Support Recovery Measurement]
  We apply Matthew's Correlation Coefficient (MCC)
  \cite{matthews1975comparison} as a statistic for measuring support
  recovery performance.  This statistic synthesizes the confusion
  matrix into a single score appropriate for unbalanced labels and is
  calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
  performance, $0$ being the performance of random guessing, and $-1$
  being perfectly opposed.
\end{remark}

\begin{remark}[Error Measurement]
  We we estimate the 1-step ahead prediction error by forming the variance matrix estimate

  \begin{equation*}
    \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
  \end{equation*}

  on a long stream of out-of-sample data.  We then report the quantity

  \begin{equation*}
    \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
  \end{equation*}

  where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
\end{remark}

\begin{figure}
  \centering
  \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006}}
  \label{fig:simulation_results}

@@

  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
    estimation.  We make comparisons against both the MCC and the
    relative log mean-squared prediction error
    $\frac{\ln\tr \widehat{\Sigma}_v}{\ln\tr \Sigma_v}$.  Results
    in Figure \ref{fig:simulation_results_comparison1} are for systems
    guaranteed to satisfy the assumptions required for Theorem
    \ref{thm:scg_recovery}.  Figure
    \ref{fig:simulation_results_comparison2} provides results for
    systems which do not guarantee the assumptions of Theorem
    \ref{thm:scg_recovery}, though the graph has a similar level of
    sparsity.}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{Fixed $T$, increasing $n$ (SCG)}
    \label{fig:simulation_results_scaling}
    \includegraphics[width=\linewidth]{new_increasing_n_simulation.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{MCC Comparison for $T \le 100$}
    \label{fig:small_T_comparison}
    \includegraphics[width=\linewidth]{new_mcc_comparison000.pdf}
  \end{subfigure}

  {\scriptsize Figure \ref{fig:simulation_results_scaling} depicts
    support recovery performance as the number of nodes $n$ increases,
    and the edge proportion as well as the number of samples $T$ is
    held fixed.  Remarkably, performance appears to degrade slowly in
    $n$ -- it is the graph topology (SCG or non-SCG) as well as the
    level of sparsity (measured by $q$) which is the main factor for
    support recovery performance.

    Figure \ref{fig:small_T_comparison} provides a support recovery
    comparison for very small values of $T$ typical of some
    applications.}
\end{figure}

\begin{figure}
  \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
  \label{fig:simulation_results_dense}
  \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}

  {\scriptsize Figure \ref{fig:simulation_results_dense} provides a
    comparison between PWGC and AdaLASSO as the density of graph edges
    (as measured by $q$) increases.  For reference,
    $\frac{2}{n} = 0.04$ has approximately the same level of sparsity
    as the SCGs we simulated.  As $q$ increases, the AdaLASSO
    outperforms PWGC as measured by the MCC.  However, PWGC maintains
    superior performance for 1-step-ahead prediction.  We speculate
    that this is a result of fitting the sparsity pattern recovered by
    PWGC via OLS which directly seeks to optimize this metric, whereas
    the LASSO is encumbered by the sparsity inducing penalty.}
\end{figure}
